---
title: "Weak Rope Analysis"
author: "Jess Gorzo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2, quietly = TRUE, verbose=FALSE)
library(stringr)
library(leaps)
library(car)
library(MuMIn)
library(emmeans)
library(interactions)
```

## Introduction

The purpose of this analysis was to compare gill nets with the number of weak links
required by NMFS (control group) to those with more than the required number (experimental group),
to determine how catch is affected by the number of weak links in a net. There were
4 experimental treatments for types of weak link configurations:

* Single End Line (One Vert line)
* Weak endline
* Weak endline and Weak Horizontal
* Weak Horizontal Line

## Methods

The relationship of catch to the experimental parameters of interest was modeled 
as an analysis of covariance (ANCOVA). This framework allows for accommodating
co-variates that can affect catch (in this scenario, nuisance variables)
while quantifying the relationships of interest, namely the effects of whether
the net was control or experimental, the experimental treatment, and the number
of weak links. The experimental design was unbalanced, though, due to the nature of depending
on volunteer participation rather than being a researcher-led endeavor with gear
designed solely for the experiment. An ANCOVA is most powerful across groups when the experimental 
design is balanced, so balancing groups in the data to the extent possible was a priority in preparing the data
for analysis. For instance, each record was assigned to a single target species, 
even if multiple were listed. I attempted to balance the target groups by
assigning records to a given target species based on what seemed to be the primary target, and/or
attempting to achieve as equivalent sizes of target species groups as possible and reasonable where multiple
species were listed.

The following target species were considered, and lumped into categories where necessary:
* bluefish
* dogfish (species lumped, since species wasn't always specified)
* menhaden
* monkfish
* shark (species lumped, since species wasn't always specified)
* skate (species lumped, since species wasn't always specified)
* Spanish mackerel

```{r cars}
load(file.path(Sys.getenv("FILEPATH"),"data/weakrope/weakrope_data.RData"))
merged <- merge(haul, string, by=c("Name","stringid"))
merged <- merged[!is.na(merged$`Target Species`),]
merged$target <- "bluefish"
merged$target[merged$`Target Species`=="butterfish"] <- "butterfish"
merged$target[merged$`Target Species`=="croaker"] <- "croaker"
merged$target[merged$`Target Species` %in% c("dogfish", "small dogfish",
                                             "smooth dogfish",
                                             "smooth dogfish, bonito", 
                                             "smooth dogfish, skate",
                                             "spiny dogfish")] <- "dogfish"
merged$target[merged$`Target Species` %in% c("menhaden", "menhaden, skate")] <- "menhaden"
merged$target[merged$`Target Species` %in% c("monkfish", "skate, monkfish", "bluefish, monkfish")] <- "monkfish"
merged$target[merged$`Target Species` %in% c("shark", "spinner shark")] <- "shark"
merged$target[merged$`Target Species` %in% c("skate", "skt", "skw", "winter skate")] <- "skate"
merged$target[merged$`Target Species` %in% c("spanish mackerel")] <- "spanish mackerel"
merged$target[merged$`Target Species` %in% c("spot")] <- "spot"
merged$Treatment[is.na(merged$Treatment)] <- "control"

testdata <- merged[,names(merged)[!names(merged) %in% c("VTR#", "Expected Soak Time", 
                                            "Estimated Soak Duration", "VTR Data",
                                            "nontarget", "Set", "Haul",
                                            "Protected Species Interaction",
                                            "panel", "buoy_buoy", "Headrope Buoyancy (lb)",
                                            "footrope_buoy", "Target Species",
                                            "hours", "minutes", "Notes/design")]]
predictors <- names(testdata)[!names(testdata) %in% c("catch", "net", "Treatment",
                                    "target")]
char_cols <- sapply(testdata, is.character)
testdata[char_cols] <- lapply(testdata[char_cols], as.factor)
```

### Variable selection

Of the suite of nuisance variables available to investigate, I first 
identified aliases and chose among them, with a preference given to numeric, continuous
variables that could be best hypothesized to affect catch. To maximize the
number of net specifications considered while restricting the set to variables
that are not aliased and not co-linear, the following net specifications
were considered:

* Net Length (ft)
* Mesh Count (vertical)
* Lead line (Spool) Weight (lbs)
* No. Floats
* No. Weak Links
* Foot rope Diameter (in)
* Head rope Diameter (in)

Similarly, the following environmental factors were considered among those
available:

* set depth
* wind speed (at set and haul)
* wave length
* current (at set and haul)
* substrate
* location (lat/lon)
* sea surface temp (SST at haul)
* maximum swell
* estimated soak duration

The data set was whittled down to records that had values for all of these variables, 
leaving 786 observations. Then, these co-variates were combined into a global model, 
and all possible subsets were investigated to select a best model by AICc. Of the
3 variables of interest, preliminary analysis implied that number of weak links would 
likely be the most influential variable of interest in the model. Thus, number of weak links was also included in 
the global model used for variable selection.

```{r analysis}
model_all <- lm(catch ~ 
                wind_speed + 
                `Wave Length (ft)` + 
                current +
                Substrate +
                `Set Depth (fa)` + lat + lon + `Wind Speed (knots)` +
                `Sea Surface (f)...27` + `Current (Knots)...28` + max_swell +
                estimated_soak +
                `Net Length (ft)` +
                `Mesh Count (vertical)` +
                `Leadline (Spool) Weight (lbs)` +
                `# Floats` + 
                `# Weak Links` +
                `Footrope Diameter (in)` +
                `Headrope Diameter (in)`,
                data=testdata)

clean_data <- testdata[,which(names(testdata) %in% c("catch",
                                                     gsub("`",
                                                          "", 
                                                          names(vif(model_all)[,1])),
                                                     "target", "net", "Treatment"))]
clean_data <- na.omit(clean_data)
```

The final models included the following nuisance variables, in addition
to the independent variables of interest:

* No. floats
* Foot rope diameter
* Mesh count
* Net length
* Set depth
* SST at haul
* Current at set
* Wind speed at set
* Max. swell

### Model structure

The relationship of catch to the experimental conditions of interest
was modeled via interaction terms with target species, allowing the variance to 
be partitioned by target. Thus, separate models were developed with interactions of
species by the each variable of interest: experimental group (control or experimental), treatment, and number of weak links. 

```{r models}
best_model_net <- aov(catch ~ `# Floats` + `Footrope Diameter (in)` +
      `Mesh Count (vertical)` + `Net Length (ft)` + `Sea Surface (f)...27` +
      `Set Depth (fa)` + current + estimated_soak + lat + max_swell + wind_speed +
      target*net, data = clean_data, na.action=na.fail)

best_model_trt <- aov(catch ~ `# Floats` + `Footrope Diameter (in)` +
      `Mesh Count (vertical)` + `Net Length (ft)` + `Sea Surface (f)...27` +
      `Set Depth (fa)` + current + estimated_soak + lat + max_swell + wind_speed +
      target*Treatment, data = clean_data, na.action=na.fail)

best_model <- aov(catch ~ `# Floats` + `# Weak Links` + `Footrope Diameter (in)` +
      `Mesh Count (vertical)` + `Net Length (ft)` + `Sea Surface (f)...27` +
      `Set Depth (fa)` + current + estimated_soak + lat + max_swell + wind_speed +
      target*`# Weak Links`, data = clean_data, na.action=na.fail)

```
## Results

While accounting for the influence of nuisance variables, whether the net
was a control or experimental net was not significant in the overall model. It 
was also not significant for any target species. Thus, the performance of a control
net did not differ significantly from the performance of an experimental net on the whole.

```{r results}
summary(best_model_net)
comp_means <- emmeans(best_model_net, ~ target * net)
pairwise_results <- pairs(comp_means, by = "target", adjust = "net")
summary(pairwise_results)
```
Considering instead the type of weak link, the treatment was also not significant in the overall model.

```{r results-trt}
summary(best_model_trt)
comp_means <- emmeans(best_model_trt, ~ target * Treatment)
pairwise_results <- pairs(comp_means, by="target")
#summary(pairwise_results)
```

On a finer scale, the model explicitly including number of weak links revealed that
the number of weak links was one of the most significant variables in this model.
However, the effect strongly differed by target species. The number of weak links 
was only a significant predictor of catch where dogfish 
and monk fish were listed as target species. Total catch for these target species 
was negatively correlated with the number of weak links in a given net. The catch
was not significantly correlated with number of weak links for any other target species.
There were not enough records where shark was the primary target to estimate a slope
for this group.

```{r results-best}
summary(best_model)
emtrends(best_model, ~ target, var="# Weak Links")
```

## Discussion

In our data set, foot rope diameter was aliased with a number of different variables describing
the net, which included tie down length, twine size, anchor weight,
manufacturer of various net components, as well as experimental parameter of the type of weak link used.
This represents a limitation of the data set, reflecting sample size and what was available given that data collection circumstances. So, we do not have enough unique combinations of these variables to properly separate them. Thus, the relationship of catch to foot rope diameter, though there was a significant linear relationship, can alternatively be interpreted as potentially 
representing other net specifications. Teasing apart those influences could be 
further investigated by testing other combinations of net measurements and
manufacturers.

Similarly, there were limited replicates of treatment per target species, and in some cases,
some treatments were never tested for a given target species. Treatment
(i.e. weak link type) was insignificant in the model, which weakens the support for
the interaction term. It is possible that clearer relationships between
treatment and catch separated by target species might be elucidated by more replicates per group,
and also a more complete experimental design to ensure each treatment is tested
for each target.

Nonetheless, the variable quantifying the differences between the control and experimental 
nets, namely the number of weak links, was highly significant in the model, and 
negatively affected the catch for 2 groups of target species.