---
title: "Weak Rope Analysis"
author: "Jess Gorzo"
date: "`r Sys.Date()`"
bibliography: references.bib
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2, quietly = TRUE, verbose=FALSE)
library(emmeans)
library(sf)
library(mapview)
library(broom)
load(file.path(Sys.getenv("FILEPATH"),
               "data/weakrope/weakrope_data_analysis.RData"))
#clean_data <- clean_data[clean_data$lon > -74.29 & clean_data$lon < -70 & clean_data$lat > 37,]
clean_data$target <- gsub("\\b([a-z])", "\\U\\1", clean_data$target, perl = TRUE)
df_sf <- st_as_sf(clean_data, coords = c("lon", "lat"), crs = 4326)

#shoot for 809 words intro
```

## Introduction

The Mid-Atlantic commercial gillnet fishery is currently in category 1 under the Marine Mammal Protection Act (MMPA), meaning it is categorically of the highest risk to marine mammals [@NOAA2024LOF]. There are currently an estimated 4020 vessels/persons participating in this fishery [@Noaa2026LOF]. Fishermen that set gillnets in the Atlantic Ocean where the floatline is above 280 fathoms deep are required to construct nets with weak links (hereby defined as a segment with a breaking strength of 1100 lbf) to allow large whales to break free when entangled [@noaa1997alwtrp]. Weak links can be constructed of rope that breaks at the required threshold, plastic, as well various as other approved materials and constructions [@noaa2015taking]. For anchored gillnets in the Mid-Atlantic brought back to port with the vessel, at least one weak link is required along the floatline per panel up to 50 fathoms long, and additional weak links every 25 fathoms for longer panels [@50cfr229]. For nets left out (i.e. not returning to port with the vessel), there must be a weak link at each flotation device and at least 5 weak links per net panel [@nmfs2014atlantic]. All panels must have a weak link placed in the center of each of the up and down lines (i.e. the lines that connect the floatline and leadline at the end of each panel). Also, each panel must have a weak link at the center of the float line, and additional weak links are required along the floatline if the panel is longer than 50 fathoms (with a maximum spacing of 25 fathoms between weak links). For the remaining 2 required weak links, there are optional configurations: either to have a weak link placed as close as possible on the floatline at each end of the panel just before the floatline meets the up and down line, or placed between the floatline tie loops between panels as well as the bridle/buoy line (depending on the net configuration).

New Jersey has not been issuing new gillnet permits offshore since 2003 [@njdep2002multispecies]. Thus, the open entry program ended in 2002, and fishermen who had existing licenses have been able to apply year after year, or transfer the license [@njac2003commercial].  Delaware Bay is exempt from these regulations [@nmfs2007alwtrp]. However, New Jersey manages gillnet fishing in the Delaware Bay via a separate permitting system [@njac2026commercial]. No new gillnet permits have been issued for Delaware Bay since 1991, and will not be issued unless the total number of permits for anchored gill nets drops below 800 [@njac7_25_18_6].

Yet, different configurations and additional weak links beyond the minimum requirement may prove increasingly beneficial for reducing by-catch of protected species. The purpose of this analysis was to compare gillnets with the number of weak links required, which served as the control group, to those with more than the required number and with novel configurations, which served as the experimental group. There were 4 experimental treatments for types of weak link configurations:

* Single end line (vertical)
* Weak end line
* Weak end line and weak horizontal
* Weak horizontal line

Further, nets were characterized by the number of weak links, to provide a continuous descriptor per string. The goal was to assess how each of these variables affected catch, broken down by target species.

## Methods

### Study Area
Coastal NJ
```{r map, include=F}
mapview(df_sf)
```

### Analysis
The relationship of catch to the experimental parameters of interest was modeled 
as an analysis of covariance (ANCOVA). This framework allows for accommodating
co-variates that can affect catch (in this scenario, nuisance variables)
while quantifying the relationships of interest, namely the effects of whether
the net was control or experimental, the experimental treatment, and the number
of weak links. The experimental design was unbalanced, though, due to the nature
of depending on volunteer participation rather than being a researcher-led
endeavor with gear designed solely for the experiment. An ANCOVA is most
powerful across groups when the experimental design is balanced, so balancing
groups in the data to the extent possible was a priority in preparing the data
for analysis. For instance, each record was assigned to a single target species, 
even if multiple were listed. I attempted to balance the target groups by
assigning records to a given target species based on what seemed to be the
primary target, and/or attempting to achieve as equivalent sizes of target
species groups as possible and reasonable where multiple species were listed.

The following target species were considered, and lumped into categories where
necessary:

* bluefish
* dogfish (species lumped, since species wasn't always specified)
* menhaden
* monkfish
* shark (species lumped, since species wasn't always specified)
* skate (species lumped, since species wasn't always specified)
* Spanish mackerel

### Variable selection

Of the suite of nuisance variables available to investigate, I first 
identified aliases and chose among them, with a preference given to numeric,
continuous variables that could be best hypothesized to affect catch. To
maximize the number of net specifications considered while restricting the set
to variables that are not aliased and not co-linear, the following net
specifications were considered:  

* Net Length (ft)
* Mesh Count (vertical)
* Lead line (Spool) Weight (lbs)
* No. Floats
* No. Weak Links
* Foot rope Diameter (in)
* Head rope Diameter (in)

Similarly, the following environmental factors were considered among those
available:  

* set depth
* wind speed (at set and haul)
* wave length
* current (at set and haul)
* substrate
* location (lat/lon)
* sea surface temp (SST at haul)
* maximum swell
* estimated soak duration

The data set was whittled down to records that had values for all of these
variables, leaving `r nrow(clean_data)` observations. Then, these co-variates were combined into a global model, and all possible subsets were investigated to select a best model by AICc. Of the 3 variables of interest, preliminary analysis implied that number of weak links would likely be the most influential variable of interest in the model. Thus, number of weak links was also included in the global model used for variable selection. The final models included the following nuisance variables, in addition to the independent variables of interest:  

* No. floats
* Foot rope diameter
* Mesh count
* Net length
* Set depth
* SST at haul
* Current at set
* Wind speed at set
* Max. swell

### Model structure

The relationship of catch to the experimental conditions of interest
was modeled via interaction terms with target species, allowing the variance to 
be partitioned by target. Thus, separate models were developed with interactions
of species by the each variable of interest: experimental group (control or
experimental), treatment, and number of weak links. 

```{r models}
best_model_net <- aov(catch ~ `# Floats` + `Footrope Diameter (in)` +
      `Mesh Count (vertical)` + `Net Length (ft)` + `Sea Surface (f)...27` +
      `Set Depth (fa)` + current + estimated_soak + lat + max_swell + wind_speed +
      target*net, data = clean_data, na.action=na.fail)

best_model_trt <- aov(catch ~ `# Floats` + `Footrope Diameter (in)` +
      `Mesh Count (vertical)` + `Net Length (ft)` + `Sea Surface (f)...27` +
      `Set Depth (fa)` + current + estimated_soak + lat + max_swell + wind_speed +
      target*Treatment, data = clean_data, na.action=na.fail)

best_model <- aov(catch ~ `# Floats` + `# Weak Links` + `Footrope Diameter (in)` +
      `Mesh Count (vertical)` + `Net Length (ft)` + `Sea Surface (f)...27` +
      `Set Depth (fa)` + current + estimated_soak + lat + max_swell + wind_speed +
      target*`# Weak Links`, data = clean_data, na.action=na.fail)

```
## Results

While accounting for the influence of nuisance variables, whether the net
was a control or experimental net was not significant in the overall model. It 
was also not significant for any target species. Thus, the performance of a
control net did not differ significantly from the performance of an experimental
net on the whole by group.

```{r results}
knitr::kable(tidy(best_model_net))
```

```{r net, include=F}
comp_means <- emmeans(best_model_net, ~ target * net)
pairwise_results <- pairs(comp_means, by = "target", adjust = "net")
summary(pairwise_results)
```
Considering instead the type of weak link, the treatment was also not significant in the overall model.

```{r results-trt}
knitr::kable(tidy(best_model_trt))
```

```{r trt, include=F}
comp_means <- emmeans(best_model_trt, ~ target * Treatment)
pairwise_results <- pairs(comp_means, by="target")
#summary(pairwise_results)
```

On a finer scale, the model explicitly including number of weak links revealed
that the number of weak links was one of the most significant variables in this
model. The interaction between the number of weak links and the target species
was also significant.

```{r results-best}
knitr::kable(tidy(best_model))
```

The effect of number of weak links strongly differed by target species. The number of
weak links was only a significant predictor of catch where dogfish and monkfish
were assigned as the representative target species for the record. Total catch for these target
species/groups was negatively correlated with the number of weak links in a given net.
To further test the robustness of this relationship, records where both bluefish and monkfish
were listed as the target species were reassigned to bluefish, and the significant
negative relationship remained. 

The catch was not significantly correlated with number of weak links for any
other target species. There were not enough records where shark was the primary
target to estimate a slope for this group.

```{r best}
knitr::kable(emtrends(best_model, ~ target, var="# Weak Links"))
```

## Discussion
For this analysis, I prioritized maximizing the number of covariates that could be considered in model selection, while removing aliased and co-linear variables from the candidate set. In our data set, foot rope diameter was significant in all models, and was aliased with a number of different variables in the data set describing the net, which included tie down length, twine size, anchor weight, manufacturer of various net components, as well as experimental
parameter of the type of weak link used. While the type of weak link used may be of interest in further analysis, this variable exhibited severe multicollinearity with several variables in the study, which may have corresponded to artifacts in the data collection process rather than a true relationship. These issues represent a limitation of the data set, reflecting sample size and what was available given that data collection circumstances. So, we do not always have enough unique combinations of the variables recorded to properly separate them. Thus, the relationship of catch to foot rope diameter, though there was a significant linear relationship, can alternatively be interpreted as potentially representing other net specifications. Teasing apart those influences could be further investigated by testing other combinations of net measurements and manufacturers.

Similarly, there were limited replicates of treatment per target species, and in
some cases, some treatments were never tested for a given target species.
Treatment (i.e. weak link type) was insignificant in the model, which weakens
the support for the interaction term. It is possible that clearer relationships
between treatment and catch separated by target species might be elucidated by
more replicates per group, and also a more complete experimental design to
ensure each treatment is tested for each target.

Nonetheless, the variable quantifying the differences between the control and
experimental nets, namely the number of weak links, was highly significant in
the model where it was considered, and negatively affected the catch for 2
groups of target species (dogfish and monkfish). Notably, the catch of the
remainder of target species/groups considered (bluefish, menhaden, shark, skate,
Spanish mackerel) were not significantly affected by the number of weak links in
this analysis. There was not enough data on sharks to model the relationship for
this group, so further research is needed to better understand how shark catch
is affected by the number of weak links in a net.

In sum, the control group did not meaningfully differ from the experimental group, but the finer scale descriptor of number of weak links proved influential for some target species groups. For target species groups not affected by the number of weak links (most of them considered in this analysis), it may be possible to use gear with more weak links that may benefit protected species while not reducing catch.

## References