---
title: "Osprey Nest Density"
author: "Jess Gorzo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(mapview)
mapviewOptions(leafletWidth = "100%", leafletHeight = 600)
library(sf)
library(sf)
library(dplyr)
library(tidyverse)
library(mapview)
library(stringr)

habitat_focal <- read_sf(file.path(Sys.getenv("FILEPATH"), "data/ospr/focal_areas.gpkg"))
```

## Opsrey Nests 
Patrick Woerner provided osprey observation data, including nest data, that he
had via the Bureau of GIS which was from largely before 2012. Kathy Clark said 
that data is more currently being kept in Osprey Watch, so I downloaded the KML
available online. This data, however, does not include the level of detail in
the attributes as what was available in-house. It would be ideal to get access
to the detailed current data I imagine is available through ENSP or Conserve Wildlife. Nonetheless,
the points here and represented in the total density are the combination of the old
and modern data.

## Opsrey Monitoring Areas and Habitat 
In lieu of spatial data delineating monitoring areas, I computed
minimum convex hulls around osprey nests within a list of focal areas
sent to us by Joe. My goal was to limit spatial extent reasonably to areas that 
are surveyed within each focal area for nests (e.g. keep areas of Monmouth Co. 
that appear to be monitored, rather than compute density on the entire
areal extent of the county). I coarsely defined
opsrey nesting habitat to be within a 1km buffer around water or a 500m wetland
buffer, informed by available life history information. I used the
Land Use/Land Cover of New Jersey 2015 dataset hosted by our NJDEP Bureau of
GIS, which is in vector format, lending well to these types of spatial data
operations. I then subset the National Land Cover Database (NLCD) 2023 to
unsuitable nesting habitat types,
namely evergreen forest, mixed forest, scrub/shrub and grassland/herbacoues land
cover categories. No nests occurred in these land cover types, which also fits
with known life history traits. Thus, I subtracted these unsuitable land cover
types from the habitat layer above within ideal distance of aquatic habitats. 
The area computed to roughly estimate density is this proxy of 
osprey nesting habitat within each minimum convex polygon.

```{r cars, echo=F}
ensp <- read_sf(file.path(Sys.getenv("FILEPATH"), "data/ospr/INPUT/ENSP_OSPR.gpkg"))
ensp_nest <- ensp[ensp$FEAT_DESC=="Nest",]

ospr <- read.csv(file.path(Sys.getenv("FILEPATH"), "data/ospr/INPUT/OspreyWatch.csv"))
ospr <- ospr %>% 
  separate(Coords, into = c("lon", "lat"), sep = ",", convert = TRUE) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326)
ospr <- st_transform(ospr, crs=st_crs(ensp_nest))

ensp_match <- ensp_nest
ensp_match$ID <- ensp_match$SF_ID 
ensp_match$Nickname <- as.character(ensp_match$LINK_ID) 
ensp_match$Substrate <- "ENSP Unrecorded"
ensp_match$Last.Watched <- str_extract(ensp_match$LAST_OBS, "\\d{4}")
ensp_match$Status <- "unknown"
ensp_match$City.County <- NA
ensp_match$State <- "New Jersey"
ensp_match$Postal.Code <- NA
ensp_match <- ensp_match[,names(ensp_match) %in% names(ospr)]
all_nests <- bind_rows(ospr, ensp_match)
#write_sf(all_nests, "ospr_nests.gpkg")

ospr_na <- ospr[ospr$Substrate %in% c("", "Other", "CM 'LB'", "Coal Dist Arm", "CM",
                              "Don't Know", "Nothing Here", "New Jersey Osprey Project",
                              "Inactive. No Nest",
                              "Other (Specify)") | is.na(ospr$Substrate),]
ospr_natural <- ospr[ospr$Substrate %in% c("Dead Tree", "Live Tree", "Marsh", "Island",
                                   "Snag In Water", "Branches/twigs",
                                   "Old Red Tail Hawk Nest",
                                   "Dead Tree W/ Platform Supports",
                                   "Ground Nest") & !is.na(ospr$Substrate),]
ospr_naturals <- rbind(ospr_na, ospr_natural)
#mapview(ensp_nest[13,]) + mapview(ospr[19,])
#mapview(all_nests, color="red") + mapview(ospr_naturals)

ospr26 <- ospr[ospr$Last.Watched=="2026",]

#load habitat_focal from focal_areas.gpkg
habitat_focal <- read_sf(file.path(Sys.getenv("FILEPATH"), "data/ospr/focal_areas_refined.gpkg"))
all_nests <- st_transform(all_nests, crs=st_crs(habitat_focal))
ospr_naturals <- st_transform(ospr_naturals, crs=st_crs(habitat_focal)) #focal_areas.gpkg

habitat_focal$n_points_nat <- lengths(st_intersects(habitat_focal, ospr_naturals))
habitat_focal$density_nat <- habitat_focal$n_points_nat / habitat_focal$area_acres

habitat_focal$n_points_all <- lengths(st_intersects(habitat_focal, all_nests))
habitat_focal$density_all <- habitat_focal$n_points_all / habitat_focal$area_acres

habitat_focal$n_points26 <- lengths(st_intersects(habitat_focal, ospr26))
habitat_focal$density26 <- habitat_focal$n_points26 / habitat_focal$area_acres

mapview(habitat_focal, zcol="NAME") + mapview(all_nests)
```

## Nest Density

```{r pressure, echo=FALSE}
sumtab <- as.data.frame(habitat_focal[,c(2:3, 7:10)])
sumtab$geom <- NULL
names(sumtab) <- c("Region", "Area (acres)", "# Nests (all years)", "Density", "# Nests (2026)", "Density (2026)")
sumtab[3,1] <- "Great Egg Harbor"
sumtab[7,1] <- "Monmouth County"
sumtab[9,1] <- "Sea Isle City"
knitr::kable(sumtab)
```
